### 2021-04-17 Chapter-1/Chapter-1.tex
Updates from Overleaf

% !TEX root = ../YourName-Dissertation.tex

\chapter{Introduction} \label{chapter1}
\section{Side-channels}
A side-channel attack is any attack that allows adversaries to infer sensitive information based on ``side-channel'' information. In 1996, Kocher published the first side-channel attack in the literature. He demonstrated that an attacker could break cryptosystems like RSA or Diffie-Hellman by measuring the execution time of a simple modular exponentiation operation. The simple modular exponentiation calculates the remainder of an exponent divided by an integer.  The simple modular exponentiation operation executes faster when the exponent bit is 0 and slower when the exponent bit is 1. An attacker can infer the whole secret key by measuring the execution time of the operation. Since then, researchers have proposed numerous side-channel attacks which exploit numerous kinds of ``side-channel'' information, including execution timing, CPU usages, power consumption,  electromagnetic radiation, and even sound.

A large portion of side-channel attacks is the joint result of the underlying computer architecture design and the above software stack.  One example is cache channel attacks. The cache is a smaller but faster memory designed to bridge the gap of the access time between the main memory and CPU registers by storing the copy of recently accessed data. Some software accesses different memory addresses when it processes different input data. Because the size of the cache is significantly smaller than the size of the main memory, the CPU can only find part of the data in the cache. Such differences can be reflected in the differences between the program execution time. Hence we have cache side-channel attacks.

Various countermeasures have been proposed to defend against side-channel attacks. Based on the above discussion, existing countermeasures can be further divided into two categories, the hardware-based method and the software-based method. Hardware-level solutions, such as reducing shared resources, adopting oblivious RAM, and using transnational memory~\cite{203878,217537,shih2017t,Zhang:2015:HDL:2775054.2694372}, need to adopt new hardware features and apply changes to today's modern complex computer systems, which are hard to adopt in reality. Therefore, a promising and universal direction is software countermeasures, detecting and eliminating side channel vulnerabilities (data flow from secrets to load addresses, and data flow from secrets to branch conditions) from existing code base. For example, when we examine those side-channel attacks, many of them exploit two kinds of code patterns: \emph{secret-dependent control-flow transfers} and \emph{secret-dependent data accesses}.
\begin{figure}[h]
    \begin{lstlisting}[xleftmargin=.32\textwidth, xrightmargin=.32\textwidth]
int secret[32];
...
while(i>0){
    r = (r * r) % n;
    if(secret[--i] == 1)
        r = (r * x) % n;   
}
\end{lstlisting}
    \caption{Secret-dependent control-flow transfers}
    \label{fig:secret:cf}
\end{figure}

Secret-dependent control-flow transfers happen when the value of secret can affect which branch the code executes. Figure~\ref{fig:secret:cf} shows an example. Depending on the value of \textsf{secret}, the code may or may not execute the code at line 6. As attackers, they can retrieve some information based on the execution control-flow.

\begin{figure}[h]
    \begin{lstlisting}[xleftmargin=.32\textwidth, xrightmargin=.32\textwidth]
static char FSb[256] = {...}
... 
uint32_t a = *RK++ ^ \ 
(FSb[(secret)) ^
(FSb[(secret >> 8)] << 8 ) ^
(FSb[(secret >>16)] << 16 ) ^
(FSb[(secret >>24)] << 24 );
...
\end{lstlisting}
    \caption{Secret-dependent data accesses}
    \label{fig:secret:da}
\end{figure}

Secret-dependent data accesses happen when the value of the secret can affect which memory cell the code reads or writes. Figure~\ref{fig:secret:da} shows an example. \textsf{FSb} is an array. Depending on the value of \textsf{secret}, the code may read or write different items in the array.

The key intuition is that the above side-channel attacks happen when a program accesses different memory addresses if the program has different sensitive inputs. As shown in Figure~\ref{fig:secret:cf} and Figure~\ref{fig:secret:da}, if a program shows different patterns in control transfers or data accesses when the program processes different sensitive inputs, the program could have side channels vulnerabilities. Various techniques can be used to exploit the above leakage site at different granularities. For example, cache channels can observe cache accesses at the level of cache sets~\cite{liu2015last}, cache lines~\cite{184415} or other granularities. Other kinds of side-channels like controlled-channel attack~\cite{7163052}, can observe the memory access at the level of memory pages. So a natural defense approach is to remove the above code patterns in the software.

However, eliminating side-channel vulnerabilities from real-world software code base is not easy. First, some vulnerable implementations actually perform better. For example, RSA implementations usually adopt the CRT optimization, which is faster but vulnerable to fault attacks~\cite{aumuller2002fault}. Fixing vulnerabilities can introduce new weaknesses. We are in a dilemma between the performance and the security. Second, it is hard to detect side-channel leakages in reality. While recent work~\cite{203878,217537,Wichelmann:2018:MFF:3274694.3274741,Brotzman19Casym,236338,182946} can detect many side-channel vulnerabilities, they either suffer from expensive computation costs or high false positives and false negatives. For example, recent tools like DATA~\cite{217537} reports 2,248 potential leakage sites for the RSA implementation in OpenSSL 1.1.0f\@. Further analysis shows 1,510 leaks can be dismissed. But that leaves 460 data-access leaks and 278 control-flow leaks. Many of these vulnerabilities have not been fixed by developers for a variety of reasons. Third, most vulnerabilities pose a negligible risk. Although some vulnerabilities result in the key being compromised~\cite{184415, aumuller2002fault} entirely, many others are less severe in reality. Therefore, we need a proper quantification metric to assess the sensitivity of side-channel vulnerabilities, so a developer can efficiently triage them.

In summary, it is hard to apply software approaches in practice due to the following reasons.

\begin{enumerate}
    \item It is hard to identify side-channel vulnerabilities in real-world software. In general, many side-channel vulnerabilities are low-level problems. Control-flows or memory accesses may not be side-channel vulnerabilities if they are independent of original secret inputs. Recent work can identify side-channel leakages. But those tools can only work on small programs or need to apply some domain knowledge of the target program to simplify the analysis.
    \item Side-channel vulnerabilities are ubiquitous. While some vulnerabilities can result in the key being entirely compromised, many other vulnerabilities are hard to exploit. Moreover, some vulnerable implementations perform better. Recent work can identify thousands of leaked points, but most of them are not patched by developers because those leakages are either false positives or leaked very little information.
\end{enumerate}

In this dissertation, we try to address the above problem from the following two aspects.
\begin{enumerate}
    \item \textbf{Detection.} We propose a fast and precise side-channel detection method. Different from previous trace comparison based methods, our method has fewer false positives and false negatives. Compared with methods based on symbolic execution and abstract interpretation, our methods can work on real-world cryptography libraries. Symbolic execution provides fine-grained information, but it is expensive to compute. Therefore, prior symbolic execution work~\cite{203878,236338,Brotzman19Casym} either analyzes only small programs or applies domain knowledge~\cite{203878} to simplify the analysis. We examine the bottleneck of the trace-oriented symbolic execution and optimize it to work on real-world applications.
    \item \textbf{Quantification.} We propose methods to quantify the amount of information leakage precisely that an attacker can retrieve during the side-channel attacks. We define the amount of leaked information as the cardinality of possible secrets based on an attacker's observation. Before an attack, an adversary has a large but finite input space. Whenever the adversary observes a leakage site, they can eliminate some impossible inputs and reduce the input space's size. In an extreme case, if the input space's size reduces to one, an adversary has determined the input, which means all secret information (e.g., the entire secret key) is leaked. By counting the number of inputs~\cite{10.1007/11499107_24}, we can quantify the information leakage precisely. 
\end{enumerate}

\section{Contribution}
This dissertation makes the following contributions. First, we propose a novel method that can detect side-channel vulnerabilities in real-world applications fast and precisely. Second, we propose methods that can quantify each side-channel leakages precisely. In addition, we study two kinds of threat models, single-trace attacks, and multiple-trace attacks. In summary, we make the following contributions:

To address Problem 1, we make the following contributions on detecting side-channel leakages:

\begin{itemize}
    \item We designs a symbolic analysis method that can detect secret-dependent control-flow transfers and secret-dependent data access. The key idea is to only symbolically execute those machine instructions that are relevant to vulnerabilities while executing other instructions natively. We implement the above method in a tool called \detect{}. The evaluation results shows that \detect{} is 3 to 100 times faster than existing tools while identifying all vulnerabilities when we run it on the same library.  We do not need to apply domain knowledge to only part of the original program, which simplify the analysis. In addition, \detect{} yields new leakages which are not identified by previous tools. We confirm those findings with developers.
\end{itemize}

To address Problem 2, we make the following contributions on quantifying side-channel leakages in a single trace attack:

\begin{itemize}
    \item We propose a method that can quantify fine-grained leaked information from side-channel vulnerabilities that result from actual attack scenarios in a single trace attack. Our approach differs from previous ones in that we precisely calculate the amount of leakages. We model the information quantification problem as a counting problem and use a Monte Carlo sampling method to estimate the information leakage.
    \item We implement the method into a tool and apply it to several pieces of real software. \tool{} successfully identifies previous unknown and known side-channel vulnerabilities and calculates the corresponding information leakage. Our results are useful in practice. The leakage estimates and the corresponding inputs that triggers the vulnerability can help developers to triage and fix the vulnerabilities.
\end{itemize}

To address Problem 2, we make the following contributions on quantifying side-channel leakages on a multiple trace attack:
\begin{itemize}
    \item We propose a novel method that can detect and analyze the effect of each side-channel leakage sites automatically. Our analysis combine the information from both the source code and the run-time execution, which makes the method more effective in finding the leakages. By running the chi-squared test, the method can find independent leakages. Attacker can combine multiple leakage sites to retrieve more information as well.
    \item We analyze the amount of information leakage from side-channel vulnerabilities based on the channel capacity. We propose a method that can estimate the lower bound of information leakages. The theory can serve as the foundation of the future work on quantify the information leakage based on the fuzzing.
    \item We implement the above method in a tool called \ctool{} and evaluate the tool with several benchmarks and several real-world applications such as OpenSSL, mbedTLS, Libgcrpyt, TinyDNN, and GTK. The results show that our tool can detect and quantify the side-channel leakages effectively. Moreover, the leakage reports given by the \ctool{} can help developers fix the reported vulnerabilities.
\end{itemize}

\section{Dissertation Organization}
The rest of the dissertation is organized as follows. We present background and related work of the dissertation in Chapter~\ref{chapter2}. Chapter~\ref{chapter3} presents a novel method to detect side-channel vulnerabilities in binary executables. Chapter~\ref{chapter4} presents a method that can quantify each side-channel leakage in a single trace attack. Chapter~\ref{chapter5} extends the approach and quantifies the side-channel leakage in a multiple-trace attack. Chapter~\ref{chapter6} concludes the dissertation and proposes the future work.