### 2021-04-17 SupplementaryMaterial/Acknowledgments.tex

% Place acknowledgments below.
I am tremendously fortunate to have been advised by Dr. Dinghao Wu.  I thank Dinghao for inviting me to EPFL on his sabbatical. A large portion of the research presented in the dissertation was finished when I worked in Switzerland. The knowledge and skills I learned from him will also help me make progress in my future career. I also thank him for treating me as an independent researcher since the fourth year of my Ph.D. study. I had a very hard time during the pandemic. Without his encouragement and support, I could not have finished the program.

Besides my advisor, I also thank my committee members, Dr. Peng Liu, Dr. Sencun Zhu, Dr. Benjamin Hanrahan, and Dr. James Larus, for their time, reviewing my dissertation. Without the strong encouragement from Dr. Peng Liu's former students in 2015, I would not have applied for Penn State.

Pursuing a Ph.D. can be extremely stressful, and summer internships are always the relaxing and productive moments during my study. I would like to thank Dr. Kang Li, Mr. Zhaofeng Chen, Dr. Mingshen Sun, Dr. Greg Zaverucha, Dr. Kevin Kane, and Dr. Brian LaMacchia for their help during my internships at Baidu X-Lab and Microsoft Research.  I also thank my friends and collaborators at Penn State University and EPFL.

This research was supported in part by the National Science Foundation (NSF) under grant CNS-1652790, and the Office of Naval Research (ONR) under grants N00014-16-1-2912, N00014-16-1-2265, and N00014-17-1-2894. Any opinions, conclusions in the paper do not reflect the views of funding agencies.