### 2021-06-01 SupplementaryMaterial/Acknowledgments.tex
Updates from Overleaf

% Place acknowledgments below.
I am tremendously fortunate to have been advised by Dinghao Wu.  I thank Dinghao for inviting me to EPFL on his sabbatical. A large portion of the research presented in the dissertation was finished when I worked in Switzerland. I also thank him for treating me as an independent researcher during my Ph.D. study. I had a very hard time during the pandemic. Without his encouragement and support, I could not have finished the program.

Besides my advisor, I thank my committee members, Peng Liu, Sencun Zhu, Benjamin Hanrahan, and James Larus, for their time, reviewing my dissertation.

Pursuing a Ph.D. can be extremely stressful, and summer internships are always the relaxing and productive moments during my study. I would like to thank Kang Li, Zhaofeng Chen, Mingshen Sun, Greg Zaverucha, Kevin Kane, and Brian LaMacchia for their help during my internships at Baidu X-Lab and Microsoft Research.  I also thank my friends and collaborators at Penn State University and EPFL.

This research was supported in part by the National Science Foundation (NSF) under grant CNS-1652790, and the Office of Naval Research (ONR) under grants N00014-16-1-2912, N00014-16-1-2265, and N00014-17-1-2894. Any opinions, or conclusions in the paper do not reflect the views of funding agencies.