### 2021-06-01 response.txt
Updates from Overleaf

Dear xxx,

Thank you for reviewing my dissertation. I have gone through the details of each
comment and revised my dissertation multiple rounds to eliminate any logic flows and
fix all the typos and grammar errors. In summary, the revision covers the following aspects.

1. Fuzzing – Why did you use fuzzing?

Abacus identifies and quantifies side-channel leakages in a single trace. So it
suffers from the problem of code coverage.  That is, Abacus misses side-channel
leakages in the unexecuted code path.

I have a few choices to address the limitation. First, I can use static methods
(verification or program analysis). Although I am interested in static analysis, I am
not a big fan of current static analysis tools, especially for side-channel research.
There are many papers on static side-channel analysis. Nearly all of these tools are
used to prove the absence of side-channel leakages. However, most of them have high
false-positive rates.

On the other hand, there are many success stories with dynamic tools (e.g., SAGE,
AFL, Valgrind) in recent years. In particular, fuzzing is becoming a hot trend these
years, especially after DARPA CGC in 2016. The fuzzing method is suitable for a task
such as side-channel detection. If a program executes two different code paths or
accesses two different variables with two different input values, the program is
vulnerable to side-channel attacks. When we report side-channel leakages sites to
developers, we can tell them the two inputs. Developers can quickly reproduce the
leakage sites with the inputs.  Interestingly, there are a few (not many) related
papers. That's why I used fuzzing at the beginning of the project.

2. Discuss the limitations of your research e.g., can only handle address based …
you might want to include some discussion on the points (3), (4), and (5) below.
The research presented in the dissertation has a few limitations.

a. The methods in my dissertation can not find all kinds of side-channel leakages. No
tool can find all forms of leakages. Even for address-based side-channel leakages,
Abacus and Quincunx can not find all the leakages due to the code coverage problem.
Some verification tools can prove the absence of side-channel leakages for some small
code snippets. However, to the best of my knowledge, no tools as yet can prove the
absence of side-channel leakages in production software.

b. The quantification result only applies to specific threat models. For example,
Abacus quantifies the number of leaked bits during one real execution. So it is
possible that one leakage site leaks little information in one execution but leaks
much information in another execution.

c. The research in the dissertation can only handle address-based side-channel
attacks. Many side-channel attacks infer secret data based on other side-channel
signals (e.g, timing, EM signals). However, the root cause of these attacks is still
the same. That is, the program accesses different addresses when it processes
different input secrets. Under the circumstance, our methods can still detect these
leakage sites.

3. Can you extend your research (especially the quantification part) to include other
side-channel signals?

It is possible to extend the quantification part to include other side-channel
signals (e.g., Sound, CPU usages, EM signals). The key is to model the noises
properly. In my dissertation, we assume that an attacker can have a noise-free
observation. While a noise-free observation is possible for some address-based
side-channel attacks, the assumption is too ideal for many other side-channel
attacks. The good news is that there are some mature methods in information theory to
model noises (e.g., Gaussian noise). The noisy-channel coding theorem provides a
computable method to estimate the information that can be reliably transferred over a
channel between secret data and attackers' observations. From a research project
aspect, I hope to have a stronger motivation before starting the project.

4.  bits vs. Q bits. Lower bounds?       Improve the presentation
       Make the comparison more meaningful
       Etc.

It is hard to explain Entropy. There is a rumor about entropy between John von
Neumann and Claude Shannon. When Shannon developed the concept of information theory,
he did not have a good idea to name the measurement of uncertainty in a system.
Neumann told Shannon, ``You should call it entropy, for two reasons. In the first
place, your uncertainty function has been used in statistical mechanics under that
name, so it already has a name. In the second place, and more important, nobody knows
what entropy really is, so in a debate, you will always have the advantage.'' During
my dissertation proposal, everyone does not really understand how we quantify the
amount of leaked information in Abacus. In my final oral exam, everyone is confused
with the term ``lower bounds'' except my advisor.

a. Abacus quantifies the amount of leaked information for a given address access
behavior. It is a measure of the amount of leaked information that an attacker can
get, given that another event has already occurred.

b. Quincunx defines the amount of leaked information by the empirical entropy of the
observations. For a determinism program, channel capacity is the tight upper bound of
the empirical entropy of the observation. In the early draft of the paper, I use the
term ``conservative''. Dr. Wu thinks ``conservative'' might be a confusing term and
suggests me with the term ``lower bound''. That's the reason that I used the term
``lower bound'' during my PhD defense. I have removed the term ``lower bound'' to
avoid any confusion in my dissertation.

c. I agree with Dr. Liu's comments that the quantification metrics in Chapter 4 and
Chapter 5 are different. The reason that I compare them together in Table 5.3 is that
they have the same dimensional unit (bit) under the dimensional analysis. I have
removed the comparison from the evaluation section in the current version.

To address the above concerns, I have made the following revisions.

a. I revised the description in Chapter 5. I avoided the use of term ``lower bound''
and tune my presentation.

b. I revised Table 5.3. The evaluation section does not compare Quincunx with Abacus
in the current version because they are calculated with different methods and have
different threat models.

5. iid assumption – does it fit your research on side-channel?

It is true that Quincunx requires the input variables to be independent and
identically distributed. That's why I applied a blackbox fuzzing instead of a
coverage-guided fuzzing. As long as each time we get a random input from the whole
input space (Populations) that is independent of the previous input, the input
variables should satisfy the i.i.d. assumption. For AES-128 encryption, each time we
randomly generate a 128 bit key from $0$ to $2^{128}$, the sequence of key values
follows the i.i.d. assumption.

The situation is a little different for a machine learning library. In my
dissertation, we also evaluated Quincunx on a deep learning model. Each time, we
randomly select an item from a dataset (Samples). If the dataset is biased, then we
can not get correct results. The good news is that the most well-known data in
machine learning is i.i.d. i.i.d. is the foundation of supervised machine learning
methods. The assumption of i.i.d. simplify many methods by assuming the data won't
change over the time. In my dissertation, we used MINIST, a well-known handwriting
digit dataset. We believe the i.i.d. assumption holds under the circumstance.

I have revised my dissertation to include the discussion on i.i.d.

6. Some feedback I gave before your defense.

As suggested by Dr. Wu, I added a new chapter on Discussion and Limitations (Chapter 6),
and discussed potential ways and related work to the mitigated reported side
(Chapter 2).

7. Improve Grammar.

Dr. Larus helped me a lot in fixing grammar issues. I also revised the dissertation
multiple times. In addition, I used Grammarly to proofread my dissertation.

I have incorporated some discussions above into my dissertation. Please let me know if you
have any feedback. Thank you for reviewing my dissertation again.

Thanks,
Qinkun