### 2021-06-01 SupplementaryMaterial/Acknowledgments.tex
Updates from Overleaf

% Place acknowledgments below.
I am tremendously fortunate to have been advised by Dinghao Wu.  I thank Dinghao for inviting me to EPFL on his sabbatical. A large portion of the research presented in the dissertation was finished when I worked in Switzerland. I also thank him for treating me as an independent researcher during my Ph.D. study. I had a very hard time during the pandemic. Without his encouragement and support, I could not have finished the program.

The members of my committee, Peng Liu, Sencun Zhu, Benjamin Hanrahan, and James Larus, provided helpful comments on my dissertation. Prof. Liu and Prof. Zhu greatly improved this dissertation with their comments from a security research angle. Prof. Hanrahan helped me improve the grammar of my dissertation proposal and my dissertation. Prof. Larus is a senior researcher and my last English teacher. I learned a lot on the written English from him.  His comments are also very close to what I get from reviewers later. However, I did not take his comments seriously at the very beginning.

Pursuing a Ph.D. can be extremely stressful, and summer internships are always the relaxing and productive moments during my study. I would like to thank Kang Li, Zhaofeng Chen, Mingshen Sun, Greg Zaverucha, Kevin Kane, and Brian LaMacchia for their help during my internships at Baidu X-Lab and Microsoft Research.  I also thank my friends and collaborators at Penn State University and EPFL.

Finally, I would like to thank my parents for their continuous emotional and financial support. My parents always hope that I can study in a university that is near to them. However, they still expressed their full support for me when I decided to come to U.S. as a graduate student.

This research was supported in part by the National Science Foundation (NSF) under grant CNS-1652790, and the Office of Naval Research (ONR) under grants N00014-16-1-2912, N00014-16-1-2265, and N00014-17-1-2894. Any opinions, or conclusions in the paper do not reflect the views of funding agencies.