### 2021-06-16 SupplementaryMaterial/Vita.tex
Updates from Overleaf

% Place your vita below.

Qinkun Bao finished his Ph.D. in the College of Information Sciences
and Technology, The Pennsylvania State University. He was a member of the Software 
System Security Research Lab, where he worked under the supervision of Dr. Dinghao Wu.
His research interests span reverse engineering, binary analysis, and trusted computing. Before he went to Penn State, he received his bachelor's degree in Electronic Engineering and Information Science (EEIS) from University of Science and Technology of China (USTC) in 2016.